---
title: "lab_1"
author: "od"
date: "12/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
```


1. Estimate the letter occurrence probabilities of all 26 letters in English by measuring a paragraph of English text from Wikipedia. (hint use strsplit() to split a paragraph into individual letters) (1 point).

```{r}
wp <- "A complete blood count (CBC) is a set of medical laboratory tests that provide information about the cells in a person's blood. The CBC indicates the counts of white blood cells, red blood cells and platelets, the concentration of hemoglobin, and the hematocrit (the volume percentage of red blood cells). The CBC is often carried out as part of a medical assessment, and can be used to monitor health or diagnose diseases. The results are interpreted by comparing them to reference ranges, which vary with gender and age. A landmark in the automation of blood cell counts was the Coulter principle (1953), which uses electrical impedance measurements to count blood cells and measure their sizes. This technology remains in use in many automated analyzers. Further research in the 1970s involved the use of optical measurements to count and identify cells, which enabled the automation of the white blood cell differential. "
#strsplit(wp, split = "")
let1 <- str_split(wp, pattern = "")
class(let1)
table(let1)

let2 <- str_split(wp, pattern = "") %>% unlist() %>% str_to_lower()
class(let2)
table(let2)
letter_count <- as_tibble(table(let2))

### removing space and punctuation
clean_count <- letter_count %>% 
  filter(let2 %in% letters)
clean_tab <- clean_count %>% 
  rename(freq = n) %>% 
  mutate(pct = freq / sum(freq)) %>% 
  arrange(desc(freq))

clean_tab %>% knitr::kable()

```


2. Generate “random” strings of letters that are sampled from a distribution where letter occurrence probability is the same as natural English. Use the probabilities for each letter from this wikipedia article, or use your own estimates from the previous question (2 points).

```{r}
let_sample <- sample(x = clean_tab$let2, size = 5, 
                     replace = TRUE, prob = clean_tab$pct)
let_sample

my_letters <- sample(x = clean_tab$let2, size = 50*5, 
                         replace = T, prob = clean_tab$pct)
mat2 <- matrix(my_letters, ncol = 5)
paste0(mat2[1,], collapse = "")
dim(mat2)[1]

random_strings <- c()
for (i in 1:dim(mat2)[1]){
  random_strings[i] <- paste0(mat2[i,], collapse = "")
}
random_strings
```

3. Generate a random walk of 10,000 steps. In a random walk, you are simulating the process of randomly taking a step up or down, as if you are on an infinite staircase. At each step you flip a coin. If you get heads you go up one step, if you get tails you go down one step. Start on step 0, then simulate a random walk for 10,000 steps. Your vector should preserve the step number for each step. For example, if the the first three steps were all heads, then the vector would begin with 0,1,2,3, which indicates a single step up each time. Plot the first 1,000 steps. (1 point)

```{r}
step <- c(0)
for (i in 1:10000){
  coin_flip <- sample(c(1, -1), 1)
  step[i+1] <- step[i] + coin_flip
}
coin_flip
```


```{r}
step %>% head
step %>% tail
```

```{r}
plot(step[1:1000])
```


4. What was the most positive and most negative step reached out of 10,000? (1 point)

```{r}
max(step)
min(step)
summary(step)
hist(step)
```


5. What was the longest run of steps where all steps were positive numbers. For example, in the sequence: 1,2,3,2,1,0,-1,-2,-1,-2,-1,0,1,2,3; the answer is 5 because the first five values were all positive, and this was the longest sequence of positive values. (1 point).

```{r}
step > 1
logical_step <- step %>% sign

sequence <- c()
counter <- 0
c(sequence, counter)

for (i in 1:length(logical_step)){
  if(logical_step[i] == 0){
    sequence <- c(sequence, counter)
    counter <- 0
  } else{
    counter <- counter + logical_step[i]
  }
}

sequence %>% summary()

```

