---
title: "Sampling Distributions"
author: "od"
date: "12/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
library(tidyverse)
theme_set(theme_light())
```

## 1. Probability Dist

```{r}
dist1 <- rnorm(1000)
hist(dist1)
lattice::histogram(dist1)
ggplot(aes(dist1),data = as_tibble(dist1))+
  geom_histogram(color = 'white', bins = 100)

dist2 <- as_tibble(rnorm(10000, mean = 0, sd = 1))
dist2 %>% 
  ggplot(aes(value))+
  geom_histogram(color = 'white', bins = 100)
```


#### How often did we sample a value larger than 2.5? 
```{r}
sum(dist1>2.5)
```

#### What is the probability that chance would produce a value larger than 2.5? 
```{r}
sum(dist1 > 2.5)/length(dist1)
```

#### What about any other value for this distribution?
```{r}
sum(dist1> 1.5)/length(dist1)
```

### Prob density function: plotting a normal curve

```{r}
sum(dist1 < 2.5)/length(dist1)
dnorm(2.5)
```

```{r}
dnorm(-4:4, mean = 0, sd = 1)
pdf <- dnorm(-4:4, mean = 0, sd = 1)
plot(pdf)
data_points <-  tibble(values =  dnorm(-4:4, mean = 0, sd = 1), x = -4:4)
data_points %>%
  ggplot(aes(x, values)) +
  geom_point()
```

```{r}
data_line <-  tibble(values =  dnorm(seq(-4, 4, .001), mean = 0, sd = 1), 
                     x = seq(-4, 4, .001))
data_color1 <- data_line %>% filter(x > 2.5)
data_line %>%
  ggplot(aes(x, values)) +
  geom_line()+
  geom_ribbon(data = data_color1,
              fill = 'red',
              aes(ymin = 0, ymax = values))
```

### Prob density function: getting p-value

```{r}
pnorm(2.5, 0, 1) 
pnorm(2.5, 0, 1, lower.tail = FALSE)
```

```{r}
data_color2 <- data_line %>% filter(x < 2.5)
data_line %>%
  ggplot(aes(x, values)) +
  geom_line()+
  geom_ribbon(data = data_color2,
              fill = 'wheat',
              aes(ymin = 0, ymax = values))
```

### Prob density function: getting z-value
```{r}
qnorm(0.006209665)
qnorm(.05) 
qnorm(.05, lower.tail = FALSE)
```


## 2. Sampling Distribution

```{r}
sample_mean <- mean(rnorm(10, 0, 1))
sample_mean
```


#### Plotting sample means distribution
```{r}
sm1 <- replicate(1000, mean(rnorm(10, 0, 1)))
sm1 %>% lattice::histogram()
sm1 %>% lattice::densityplot()
```

#### Standard Error of means: $\text{SEM} = \frac{\sigma}{\sqrt{N}}$
```{r}
sd = 1
n = 10
sd / sqrt(n)
sem1 <- sd(sm1)
sem1
```


## 3. Assignment

#### Verification
```{r}
pnorm(2.5, 0, 1, lower.tail = F)
```

```{r}
m <- rnorm(1000000, 0, 1)
slt <- sum(m > 2.5)
slt / 1000000
```

#### m_score = 100 sd_score = 10  N=10000  Prob score  >3 sd
```{r}
m_score <- 100
sd_score <- 10
q_3sd <- 130
N_score <- 10000

p_score <- pnorm(q = q_3sd, mean = m_score, sd = sd_score, lower.tail = F)

round(p_score * N_score)
```

```{r}
obs <- rnorm(10000, 100, 10)
sum(obs > 130)/10000 
```

#### m_score = 10 sd_score = 20  N=25  Prob mean >= 12
```{r}
samp_dist <- replicate(10000, mean(rnorm(25, 10, 20)))
hist(samp_dist)
sum(samp_dist>= 12) / 10000
```

#### m_score = 10 sd_score = 20  N= 100  Prob mean >= 12
```{r}
samp_dist <- replicate(10000, mean(rnorm(100, 10, 20)))
hist(samp_dist)
sum(samp_dist>= 12) / 10000
```

#### m_score = 10 sd_score = 20  N= 25  Prob sd <= 15
```{r}
samp_dist <- replicate(10000, sd(rnorm(25, 10, 20)))
hist(samp_dist)
sum(samp_dist <= 15) / 10000
```

